Increased Education Level of Women Associates with Higher Occupational Prestige and Happiness

### 2023-03-16 outputs/paper/paper.qmd
Fixed Graph Error

```{r}
#| label: fig-1
#| fig-cap: displays that the overall increase in the mean education level percentage of women over 20 years from 2000 to 2021 is from 13.0 to 13.7.
#| echo: false
#| warning: false

library(tidyverse)

@@

merged_gss |> group_by(Year) |> summarize(mean= mean(Highest_Educ_Level, na.rm = TRUE)) |>
  ggplot(aes(x = Year, y = mean)) +
  geom_line(group = 1) +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.9, hjust=1)) +
  labs(
    title = "Mean highest education level of Women over 20 Years",
    x = "Year",
    y = "Mean Educ Level %"
  ) 
```

@@

```{r}
#| label: fig-2
#| fig-cap: caption goes here
#| echo: false
#| warning: false

@@

merged_gss |> group_by(Year) |> summarize(mean = mean(Occupational_Prestige_Pct, na.rm = TRUE)) |>
  ggplot(aes(x = Year, y = mean)) +
  geom_line(group = 1) +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.9, hjust=1)) +
  labs(
    title = "Occupational Prestige of Women over 20 Years",
    x = "Year",
    y = "Mean Occ. Prest. %"
  ) 
```

@@

```{r}
#| label: fig-3
#| fig-cap: caption goes here
#| echo: false
#| warning: false

library(tidyverse)

merged_gss |> group_by(Year) |> filter(!is.na(Happiness)) |> count(Happiness) |>
  ggplot(aes(x = Year, y = n, fill = Happiness)) +
  geom_col(position = "dodge") +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.9, hjust=1)) +
  labs(
    title = "Womens Happiness over 20 Years",
    x = "Year",
    y = "Count"
  )
```

In @fig-3, the counts of \"Very Happy\" respondents peaked in 2000, 2006 and 2021. In 2008, \"Very Happy\" respondents dropped but steadily increased in the years afterwards. A similar trend can be observed with \"Pretty Happy\" respondents, with peaks in 2000 and 2006, and a steady increase since 2008. There is a spike in \"Not Too Happy\" responses in 2006, where both mean education level and mean occupational prestige dropped. However, where there is another spike in \"Not Too Happy\" responses in 2018, there is no correlating drop in mean occupational prestige and education level. Figures Y and Z show steady growth in their respective display values, whereas in Figure X, \"Pretty Happy\" and \"Very Happy\" counts experience similar growth but \"Not Too Happy\" counts fluctuate per year.

@@

Figures X, Y and Z show growth in their values, with Figure X specifically showing growth with the values correlated to being happy. We can reasonably deduce that there is some positive relationship between happiness, the prestige of a woman's occupation and a woman\'s peak education level.

https://www.ellevatenetwork.com/articles/9201-eight-surprising-reasons-women-are-actually-happier-at-work-than-men

From ABOVE CITATION and 2 CITATIONS ABOVE a factor for happiness in working women is the demolishment of male-favored gendered rules in the workplace. In the workforce, the old-fashioned gendered rules dictate that men are expected to succeed and women are not. However, when women succeed, those gendered rules begin to crumble and their success is seen as a victory. Over a 20 year period, working women\'s happiness may not just be correlated to their job pedigree, but also due to the collapse of the negative stigma towards women in work.

>>>>

```{r}
#| label: fig-1
#| fig-cap: A graph detailing that the overall increase in the mean highest education level percentage of women over a 20 year period from 2000 to 2021 is from 13.0 to 13.7.
#| echo: false
#| warning: false

@@

merged_gss |> group_by(Year) |> summarize(mean = mean(Highest_Educ_Level, na.rm = TRUE)) |>
  ggplot(aes(x = Year, y = mean)) +
  geom_line(group = 1) +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.9, hjust=1)) +
  labs(
    title = "Mean Highest Education Level of Women Over 20 Years",
    x = "Year",
    y = "Mean Educ Level %" 
  )
```

@@

```{r}
#| label: fig-2
#| fig-cap: A display showing the rise of the mean occupational prestige in women ranging from years 2000 - 2021.
#| echo: false
#| warning: false

@@

merged_gss |> group_by(Year) |> summarize(mean = mean(Occupational_Prestige_Pct, na.rm = TRUE)) |>
  ggplot(aes(x = Year, y = mean)) +
  geom_line(group = 1) +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.9, hjust=1)) +
  labs(
    title = "Mean Occupational Prestige of Women over 20 Years",
    x = "Year",
    y = "Mean Occ. Prest. %"
  ) 
```

@@

```{r}
#| label: fig-3
#| fig-cap: A graph of answers to the Happiness question by women ranging from the years 2000 - 2021
#| echo: false
#| warning: false

merged_gss |> group_by(Year) |> filter(!is.na(Happiness)) |> count(Happiness) |>
  ggplot(aes(x = Year, y = n, fill = Happiness)) +
  geom_col(position = "dodge") +
  theme(axis.text.x = element_text(angle = 45, vjust = 0.9, hjust=1)) +
  labs(
    title = "A Count of Womens Happiness over 20 Years",
    x = "Year",
    y = "Count"
  )
```

In @fig-3, the counts of \"Very Happy\" respondents peaked in 2000, 2006 and 2021. In 2008, \"Very Happy\" respondents dropped but steadily increased in the years afterwards. A similar trend can be observed with \"Pretty Happy\" respondents, with peaks in 2000 and 2006, and a steady increase since 2008. There is a spike in \"Not Too Happy\" responses in 2006, where both mean education level and mean occupational prestige also dropped. However, where there is another spike in \"Not Too Happy\" responses in 2018, there is no correlating drop in mean occupational prestige and education level. @fig-1 and @fig-2 show steady growth in their respective values, whereas in @fig-3 \"Pretty Happy\" and \"Very Happy\" counts experience similar growth but \"Not Too Happy\" counts fluctuate per year.

@@

@fig-1, @fig-2 and @fig-3 show growth in their values, with @fig-3 specifically showing growth with the values correlated to being happy. We can reasonably deduce that there is some positive relationship between happiness, the prestige of a woman's occupation and a woman\'s peak education level. From [@citeLinked], the work options for women have expanded due to an increase in education and societal factors. An increase in work options means that women are shifting into more professional jobs rather than clerical ones. These factors lead to an increase in women's happiness overall.

From [@citeElle], a factor for happiness in working women is the demolishment of male-favored gendered rules in the workplace. In the workforce, the old-fashioned gendered rules dictate that men are expected to succeed and women are not. However, when women succeed, those gendered rules begin to crumble and their success is seen as a victory. Over a 20 year period, working women's happiness may not just be correlated to their job pedigree, but also due to the collapse of the negative stigma towards women in work.
